Refinements for open automata

### 2021-12-03 2021-Refinement/paper/refinementarticle.tex
proof corrections
proof progess

Consider each condition of the definition of a refinement of open automata.
\begin{enumerate}
\item One has to prove it first initial configurations. By knowing that substitutions only have an effect on the variables of the open automaton they belong to, they also produce terms containing only variables of the open automaton they belong to. We have:
\begin{align*}
\mpar{\sigma_{01} \uplus \sigma_{02} \vdash R_{12}\mpar{s_{01}, s_{02}}} \wedge \mpar
{\sigma_{02} \uplus \sigma_{03} \vdash R_{23}\mpar{s_{02}, s_{03}}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{02} \uplus \sigma_{03}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}}&\implies\\ 
\underbrace{R_{12}\mpar{s_{01},s_{02}}\wedge R_{23}\mpar{s_{02}, s_{03}}}_{R_{13} \mpar{s_{01}, s_{03}}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} &\implies\\
\sigma_{01} \uplus \sigma_{03} \vdash R_{13} &\mpar{s_{01}, s_{03}}
\end{align*}

\item Consider $\mpar{s_{1}, s_{3}} \in R_{13}$ then there is a set of states $(s_{2p})^{p\in P}$ of $A_2$ relating  $s_{1}$ and $s_{3}$:
\[R_{13}(s_1,s_3)=\bigcup_{{p\in P}
}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}} \]

Consider one $s_{2p} \in (s_{2p})^{p\in P}$.

On one side, we have for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exist an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \\
			\nwedge g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline*}		
that can be re-written:
\begin{multline*}
%\everymath{\displaystyle}
\forall v \in \mpar{\fvars{\alpha_{2px}}  \cup \bigcup_{j \in J'} \!\fvars{\beta_{2pxj}}} \setminus \mpar{V_1 \uplus V_2 \uplus \fvars{t_1}}, \\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pjx} \\
			\nwedge g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} \\
	\end{multline*}	
\medskip
Adding to the two terms of the implication the predicate $R_{23}\mpar{s_{2p},s_3}$ we get:

\begin{multline*}
 \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pjx} \\
			\nwedge g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}} \land R_{23}\mpar{s_{2p},s_3}
		\end{array}}} \\
	\end{multline*}

On the other side, from the relation between $A_2$ and $A_3$ we have for any open transition $t_{2x}$ in $T_2$ originating from $s_2$:
\begin{align*}
\OTx{2}{x}{2x}{2x} \in \fOT{s_2}
\end{align*}
there exist an indexed family of OTs originating from $s_3$: 
\begin{align*}
\mpar{\OTx{3}{xy}{3xy}{3xy} \in \fOT{s_3}}^{y \in Y} \subseteq T_3
\end{align*}
such that $\forall y \in Y, J'_{2x} \cap H' = J'_{3xy} \cap H'$ and
\begin{multline*}
\everymath{\displaystyle}
V_2 \uplus V_3 \uplus \fvars{t_{2x}} \vdash\\\hspace{2em} \mpar{R_{23}\mpar{s_2, s_{3}} \wedge g_{2x} \implies \operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2x}=\alpha_{3xy} \wedge \bigwedge_\subbox{j \in J'_{3xy} \cap H'} \beta_{2jx}=\beta_{3jxy} \\
			\nwedge g_{3xy} \wedge R_{23}\mpar{s'_{2x}, s'_{3xy}}\psubst{\psi_{2x} \uplus \psi_{3xy}}
		\end{array}}} 
\end{multline*}
that can be re-written:
\begin{multline*}
\everymath{\displaystyle}
\forall v\in \mpar{\fvars{\alpha_{2x}}  \cup \bigcup_{j \in J'} \!\fvars{\beta_{2jx}}} \setminus \mpar{V_2 \uplus V_3 \uplus \fvars{t_{2x}}},\\ \mpar{R_{23}\mpar{s_2, s_{3}} \wedge g_{2x} \implies \operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2x}=\alpha_{3xy} \wedge \bigwedge_\subbox{j \in J'_{3xy} \cap H'} \beta_{2jx}=\beta_{3jxy} \\
			\nwedge g_{3xy} \wedge R_{23}\mpar{s'_{2x}, s'_{3xy}}\psubst{\psi_{2x} \uplus \psi_{3xy}}
		\end{array}}} 
\end{multline*}

\item For the third condition. First from the relation between $A_1$ and $A_2$ we have:
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2,\\ V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash \left(R_{12}\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
Second, from the relation between $A_2$ and $A_3$ we have:
\begin{multline*}
\forall \mpar{s_2, s_3} \in S_2 \times S_3,\\ V_2 \uplus V_3 \uplus \biguplus_\subbox{t_3 \in \fOT{s_3}} \fvars{t_3} \vdash \left(R_{23}\mpar{s_2, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \implies \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \right)
\end{multline*}
We get:

>>>>

We have to prove that the relation $R_{13}$ satisfies the three conditions of the definition of a refinement of open automata.
\begin{enumerate}
\item $R_{13}$ satisfies initial configurations. By knowing that substitutions only have an effect on the variables of the open automaton they belong to, they also produce terms containing only variables of the open automaton they belong to. We have:
\begin{align*}
\mpar{\sigma_{01} \uplus \sigma_{02} \vdash R_{12}\mpar{s_{01}, s_{02}}} \wedge \mpar
{\sigma_{02} \uplus \sigma_{03} \vdash R_{23}\mpar{s_{02}, s_{03}}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{02} \uplus \sigma_{03}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}}&\implies\\ 
\underbrace{R_{12}\mpar{s_{01},s_{02}}\wedge R_{23}\mpar{s_{02}, s_{03}}}_{R_{13} \mpar{s_{01}, s_{03}}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} %&\implies\\
 \end{align*}
Hence we get:
\[\sigma_{01} \uplus \sigma_{03} \vdash R_{13} \mpar{s_{01}, s_{03}}\]
\item Consider $\mpar{s_{1}, s_{3}} \in R_{13}$ then there is a set of states $(s_{2p})^{p\in P}$ of $A_2$ relating  $s_{1}$ and $s_{3}$:
\[R_{13}(s_1,s_3)=\bigvee_{{p\in P}
}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}} \]

Let's consider any $s_{2p} \in (s_{2p})^{p\in P}$.

On one side, we have for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exist an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \\
			\nwedge g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline*}		
that can be re-written:
\begin{multline*}
%\everymath{\displaystyle}
\forall v \in \mpar{\fvars{\alpha_{2px}}  \cup \bigcup_{j \in J'_{2px}} \!\fvars{\beta_{2pxj}}} \setminus \mpar{V_1 \uplus V_2 \uplus \fvars{t_1}}, \\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \\
			\nwedge g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} \\
	\end{multline*}	
\medskip
Adding to the two terms of the implication the predicate $R_{23}\mpar{s_{2p},s_3}$ we get:
\begin{align*}
R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies\hspace{5.5cm}\\ \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \land\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}} \land R_{23}\mpar{s_{2p},s_3}
		\end{array}} \\
\end{align*}

On the other side, according the relation between $A_2$ and $A_3$ we have for any open transition $t_{2px}$ in $T_2$ originating from $s_{2p}$:
\begin{align*}
\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}
\end{align*}
there exist an indexed family of OTs originating from $s_3$: 
\begin{align*}
\mpar{\OTx{3}{pxy}{3pxy}{3pxy} \in \fOT{s_3}}^{y \in Y} 
\end{align*}
such that $\forall y \in Y, J'_{2px} \cap H' = J'_{3pxy} \cap H'$ and
\begin{multline*}
\everymath{\displaystyle}
V_2 \uplus V_3 \uplus \fvars{t_{2px}} \vdash\\\hspace{2em} \mpar{R_{23}\mpar{s_{2p}, s_{3}} \wedge g_{2px} \implies \operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2px}=\alpha_{3pxy} \wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H'} \beta_{2pxj}=\beta_{3pxyj} \\
			\nwedge g_{3pxy} \wedge R_{23}\mpar{s'_{2px}, s'_{3xy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
		\end{array}}} 
\end{multline*}
that can be re-written:
\begin{multline*}
\everymath{\displaystyle}
\forall v\in \mpar{\fvars{\alpha_{2px}}  \cup \bigcup_{j \in J'_{3pxy}} \!\fvars{\beta_{2jpx}}} \setminus \mpar{V_2 \uplus V_3 \uplus \fvars{t_{2px}}},\\ \mpar{R_{23}\mpar{s_{2p}, s_{3}} \wedge g_{2px} \implies \operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2px}=\alpha_{3pxy} \wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H'} \beta_{2pxj}=\beta_{3pxyj} \\
			\nwedge g_{3pxy} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
		\end{array}}} 
\end{multline*}
The following  can be derived from both cases:
$\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$
$\forall y \in Y, J'_{2px} \cap H' = J'_{3pxy} \cap H'$ and 
implies
$\forall x \in X \forall y \in Y,  J'_1 \cap (H \cap H') = J'_{3pxy} \cap (H\cap H')$
In addition, we have

\begin{align*}
\forall v \in \mpar{\fvars{\alpha_{2px}}  \cup \bigcup_{j \in J'_{2px}} \!\fvars{\beta_{2pxj}}} \setminus \mpar{V_1 \uplus V_2 \uplus \fvars{t_1}},
\forall v\in \mpar{\fvars{\alpha_{2px}}  \cup \bigcup_{j \in J'_{3pxy}} \!\fvars{\beta_{2jpx}}} \setminus \mpar{V_2 \uplus V_3 \uplus \fvars{t_{2px}}},\\
R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies \hspace{4.5cm}\\\operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj}  \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}} \\
		\operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2px}=\alpha_{3pxy} \wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H'} \beta_{2pxj}=\beta_{3pxyj} \\
			\nwedge g_{3pxy} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
		\end{array}}	
		\end{array}} \\
\end{align*}

After simplification, we get:
\begin{align*}
R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies 
\hspace{5.5cm}\\
\operatorname*{\bigsymb{\bigvee}}_{x \in X} 
\operatorname*{\bigsymb{\bigvee}}_{y \in Y}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3pxy}\wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap (H\cap H')} \beta_{1j}=\beta_{3jpxy} \wedge g_{3pxy}\wedge \\
R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
\end{array}}
\end{align*}

Because of the domain of the substitutions of the relations, we have:
\begin{align*}
\underbrace{R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}}_{R_{13}\mpar{s_1, s_3}} \wedge g_1 \implies 
\hspace{5.5cm}\\
\operatorname*{\bigsymb{\bigvee}}_{x \in X} 
\operatorname*{\bigsymb{\bigvee}}_{y \in Y}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3pxy}\wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H'} \beta_{1j}=\beta_{3jpxy} \wedge g_{3pxy}\wedge \\
\underbrace{R_{12}\mpar{s'_1, s'_{2px}} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}}_{R_{12}\mpar{s'_1,s'_{3pxy}}}\psubst{\psi_1 \uplus \psi_{2px} \uplus \psi_{3pxy}}
\end{array}}
\end{align*}	
This allows us to conclude, with $R_{13}(s_1,s_3)=\bigvee_{{p\in P}}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}}$: 
\begin{align*}
R_{13}\mpar{s_1, s_3} \wedge g_1 \implies 
\hspace{7.5cm}\\
\operatorname*{\bigsymb{\bigvee}}_{p \in P}
\operatorname*{\bigsymb{\bigvee}}_{x \in X} 
\operatorname*{\bigsymb{\bigvee}}_{y \in Y}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3pxy}\wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H'} \beta_{1j}=\beta_{3jpxy} \\\wedge g_{3pxy}\wedge \\
{R_{12}\mpar{s'_1,s'_{3pxy}}}\psubst{\psi_1  \uplus \psi_{3pxy}}
\end{array}}
\end{align*}	
This provides the desired conclusion:
\begin{align*}
\mpar{R_{13}\mpar{s_1, s_3} \wedge g_1 \implies 
\operatorname*{\bigsymb{\bigvee}}_{z \in Z}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3z}\wedge \bigwedge_\subbox{j \in J'_{3z} \cap H'} \beta_{1j}=\beta_{3jz}\\ \wedge g_{3z}\wedge 
{R_{12}\mpar{s'_1,s'_{3z}}}\psubst{\psi_1  \uplus \psi_{3z}}
\end{array}}}
\end{align*}	
\item Concerning the condition on the deadlock reducing. According to the relation between   $A_2$ and $A_3$ we have forall $\mpar{s_{2p}, s_3} \in S_2 \times S_3$:
\begin{multline*}
 V_2 \uplus V_3 \uplus \biguplus_\subbox{t_3 \in \fOT{s_3}} \fvars{t_3} \vdash \left(R_{23}\mpar{s_{2p}, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \implies \bigvee_\subbox{t_{2p} \in \fOT{s_{2p}}} \fguard{t_{2p}} \right)
\end{multline*}
that can be rewritten:
\begin{align*}
R_{23}\mpar{s_{2p}, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \implies \bigvee_\subbox{t_{2p} \in \fOT{s_{2p}}} \fguard{t_{2p}} 
\end{align*}
By adding the term $R_{12}\mpar{s_1,s_{2p}}$ to both sides of the implication, we get:
\begin{align*}
R_{12}\mpar{s_1,s_{2p}}\wedge R_{23}\mpar{s_{2p}, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \implies R_{12}\mpar{s_1,s_{2p}}\wedge \bigvee_\subbox{t_{2p} \in \fOT{s_{2p}}} \fguard{t_{2p}} 
\end{align*}
Yet according to the relation between $A_1$ and $A_2$ we have  forall $\mpar{s_1, s_{2p}} \in S_1 \times S_2$:
\begin{multline*}
 V_1 \uplus V_2 \uplus \biguplus_\subbox{t_{2p} \in \fOT{s_{2p}}} \fvars{t_{2p}} \vdash \left(R_{12}\mpar{s_1, s_{2p}} \wedge \bigvee_\subbox{t_{2p} \in \fOT{s_{2p}}} \fguard{t_{2p}} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \right)
\end{multline*}
Then,
\begin{align*}
R_{12}\mpar{s_1, s_{2p}} \wedge \bigvee_\subbox{t_{2p} \in \fOT{s_{2p}}} \fguard{t_{2p}} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{align*}
The desired result will be obtained by transitivity of
